\pdfbookmark[1]{Acknowledgements}{acknowledgements}

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements}

Thank you:

To my advisor, Dror Bar-Natan, whose consistent meetings, guidance,
and encouragement brought me to this point, and to my supervisory committee,
Joel Kamnitzer and Eckhard Meinrenken, who helped connect me to more of the
department and its research.

To Jason Siefken and the other teaching mentors in the math
department. You have made me a better educator and communicator.

To the math department administration, especially Jemima Merisca and Sonja
Injac, whose friendly support through the manifold administrative details was
invaluable.

To Assaf Bar-Natan, Vincent Girard, Caleb Jonker, and Adriano
Pacifico, whose friendship, bike trips, and cooking sessions provided much joy;
to the people I met in the department, especially those who opened
my thesis to confirm they were included here. Thank you specifically.

To the care-taking staff at the St.\@ George campus, especially Maria
and Maria, whose diligent work and kind words have made this campus beautiful.

To the Graduate Christian Fellowship community at the U of T for
being a refreshing breadth of perspectives for a graduate student.

To Rosemary and Alan Johnstone, whose hospitality made me feel
immediately welcome in a new city.

To my family, who raised me and encouraged me in my academic pursuits from the
earliest moments: to my mom for embracing my atypical learning styles, for instilling
my love of reading, and for teaching my how to think deeply; to my dad for
giving me a love of cycling and cooking, and teaching me to notice the invisible
among us; and to my brother for all the laughter you bring to my life.

To the Perrins. You are my second family, and your support and love
for me is precious.

To my fiancée, Emily Langridge, who helped me with the motivation to
finish. I look forward to the next chapters of my life with you.

\endgroup
